Palomar 5 Gaps

### 2024-09-30 paper/main.tex
touched up gap morphology

  \subsection*{Gap Morphology}
    We study the structure of the streams that have been impact by the globular clusters. The gallery of gaps is shown in the appendix, which includes the results from all of the simulations we considered in this experiment.

    In general, how did the passage of these globular clusters effect the stream? well. A lot of this has been studied quite well and presented in Sanders et al 2016. They even consdiered different types of impactors. I.e. impactors that had different density profiles. They compared Hernquist, Plummer, NFW, and a truncated NFW. There are some minor differences between the impactors however, the conclusion is still the same. Impacts from small and dense perturbers such as globular clusters only effect the stream locally.

    Our results are consistent with those shown previously. Perturber crossings only perturb the stream locally. Global characterisitcs of the stream do not change. For instance, the stream is not warmer of thicker. Other types of interactions have been demonstrated to do this. For example THIS AUTHOR showed that XXX and do YYY. As laid out in Erkal et al 2015, perturbers pull particles from the sides towards the impact point. These particles eventually move beyond the impact point and leave an under dense center and then pile up on the sides leaving an overdensity on the shoulders.

    Interestingly, most of our gaps occur for the leading tail. 8 on the trailing tail. 21 excluding NGC~2808 are on the leading tail. We have to exclude NGC~2808 because the simulations are clearly correlated with one another, so trials are not independent. So is 8/29 consistent with a fair coin toss? In anycase, we argue this may be a coincidence since being hit should only depend on the size and location of the stream. The main difference between the tails is a $\approx$ 300-400 pc offset in $r$, yet this is negligible.

>>>>

  \subsection*{Gap Morphology}
    We analyze the structure of streams impacted by globular clusters. The gallery of gaps, found in the appendix, includes results from all the simulations considered in this experiment.

    How do globular cluster passages affect the stream? Much of this has been well studied, as discussed in Sanders et al. (2016). They even considered different types of impactors, such as those with varying density profiles (e.g., Hernquist, Plummer, NFW, and truncated NFW). While there are some minor differences between these impactors, the overall conclusion remains the same: impacts from small, dense perturbers, such as globular clusters, only affect the stream locally.

    Our results align with previous findings—perturber crossings cause local perturbations in the stream without altering its global characteristics. For instance, the stream does not become warmer or thicker, changes that have been linked to other types of interactions. For example, [AUTHOR] demonstrated that XXX can do YYY. As described by Erkal et al. (2015), perturbers pull particles from the sides toward the impact point, creating an under-dense region at the center of the gap, with overdensities piling up at the shoulders.

    Interestingly, most of the gaps we observe occur in the leading tail. However, we believe this to be coincidental. Stars escape from the globular cluster through the two Lagrange points at equal rates (cite?). The lengths of the tails are also equal, with only a small difference of approximately 400 pc in radial distance from the galactic center. This difference is negligible compared to the orbital radius of about 10 kpc. Therefore, we expect the probability of a gap appearing in the leading or trailing tail to be equal, and the presence of one gap should not influence the likelihood of another.

    Yet, the observation of only 8 gaps in the trailing tail out of 73 total gaps is inconsistent with this expectation. However, 44 of these gaps are associated with NGC2808. As shown in Fig.\ref{fig:histogram_impact_time}, all of these gaps result from the same recent encounter. We thus discard NGC~2808 for the current analysis, as its gaps violate the assumption of independent events. After excluding these, we are left with 21 gaps in the leading tail and 8 in the trailing tail across the remaining simulations. The probability of observing 8 or fewer gaps in 29 trials, assuming each has a 50\% chance, is approximately 1.2\%. While this outcome is unlikely, it is not impossible.
